### 2012-02-29 trunk/htdocs/user.tex
Some updates to tex docs
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@1514 40876441-3496-0410-a613-a2fca35b2ae2

This controls the the type of particle. Appropriate values are
\texttt{inactive}, giving a standard fully resolved LB colloid;
\texttt{active} switches on the correction to BBL appropriate
for active particles (expecting parameters $b_0$ and $b_1$ to be
set); \texttt{subgrid} gives unresolved particles (radius $< 1$
lattice unit) and no BBL.

@@

@@

\inputkey{colloid\_random\_s0} initial spin vector

@@

@@

@@

There is an important issue to understand about the (nominally)
spherical particles in LB. The input radius $a_0$ is used to
construct the link boundary conditions. However, discretisation
errors in the boundary conditions means that the `real' particle
size may be different. This real radius is the hydrodynamic radius $a_h$.
As a further complication, the hydrodynamic radius can also
be a function of viscosity for a given input radius. Some values
are given in Table~X. This values are computed using a calibration
procedure described by Nguyen and Ladd \cite{nguyen2001}.

@@

@@

May be either \texttt{ASCII} or \texttt{BINARY}. Default is \texttt{BINARY}.
Configuration I/O is always in binary. It is recommended to use
binary output for speed and disk space considerations.

@@

Individual lattice quantities can use separate I/O grids. This
may be particularly useful for the distribution data, which are
much larger than the other lattice quantities. This means a
finer decomposition may be appropriate. Clearly, the size of the
I/O decomposition cannot exceed the processor decomposition.
See the advice in section \cite{section-advice} for guidelines.

@@

\subsection{Things to Remember and Things to Avoid}
\ref{section-advice}

There are a number of simple issues to get right:
\begin{itemize}
\item
Make sure the assertions are switched off and appropriate optimisations
are switched on for production runs (see Makefile).
\item
Do not use too small a local domain size. Typically, 16$^3$ to 32$^3$
cubic local domains will give resaonble scaling. Anything smaller may
be inefficient. The best size may depend on the exact nature of the
calculation and the hardware. Note that this limitation will always
ultimately limit strong scaling.
\item
Don't have diagnsotic output (freq\_statistics) more often than is
necessary. This output requires a global communication which is slow.
\item
Use BINARY output options; ASCII output is slower, and produces larger
files.
\item
Do not use serial I/O. This will have an increasingly high overhead as
the number of processors is increases (it's probably reasonable to
128--256 MPI tasks). This is because each MPI tasks writes its data,
in turn, to a single file; this in inherently serial. To avoid this,
you need to specify the I/O decomposition for the various output
quantities of interest. This splits the system into regular blocks
(in a similar approach to the MPI decomposition), the data for which
are written to separate files in parallel. For example, if the MPI
decomposition is 8\_8\_8 (512 MPI tasks), then a reasonable I/O
decomposition may be 2\_2\_2 or 4\_4\_4. The only way to find out
what works best is to try some tests for the problem size you want to
address.
\item
Don't use the Ewald sum! This simply does not scale.
An alternative algorithm is required.
\end{itemize}

>>>>

This controls the the type of particle. Appropriate values are
\texttt{inactive}, giving a standard fully resolved LB colloid;
\texttt{active} switches on the correction to BBL appropriate
for active particles (expecting parameters $b_1$ and $b_2$ to be
set); \texttt{subgrid} gives unresolved particles (radius $< 1$
lattice unit) and no BBL.

@@

\subsubsection{The colloid state}

The full data structure for the colloid state is defined in
\texttt{src/colloid.h} and the corresponding routines which read
and write the values are in \texttt{src/colloid.c}
(note colloid singular!). Most of the state corresponds clearly
with physical properties, although there are a number of 'structural'
elements which have no physical meaning. These include the
\texttt{rebuild} flag, which instructs the code to reconstruct the
boundary links if necessary, and \texttt{deltaphi} which is involved
in accounting for conserved order parameter in the \texttt{symmetric\_lb}
free energy. Further, not all the elements are relevant is all
circumstances: e.g., the magnetic dipole is only relevant is magnetic
interactions are required, and squirmer parameters are only relevant
to active particles.

Note that it is possible in principle to have magnetic active particles,
in which case the dipole direction (\texttt{s}) and the direction of
motion vector (\texttt{m}) are allowed to be distinct.

@@

\inputkey{colloid\_random\_s0} initial spin vector (only relevant in magnetic
field)

The

@@

The squirmer parameters specify the surface boundary conditions in
an approach originating with Lighthill \cite{lighthill} and Blake
\cite{blake}. Briefly, $b_1$ sets the propulsion speed
($U = {\frac{2}{3}} b_1$), while $b_2$ sets the
particle stresslet. The ratio $\beta = b_2/b_1$ determines the
mode of the squirmer motion; see \cite{isaac} for further details.

@@

This position must be within the current system, or no particle will
be created.

@@

There is an important issue to understand about the (nominally)
spherical particles in LB. The input radius $a_0$ is used to
construct the link boundary conditions. However, discretisation
errors in the boundary conditions means that the `real' particle
size may be different. This real radius is the hydrodynamic radius $a_h$.
As a further complication, the hydrodynamic radius can also
be a function of viscosity for a given input radius. Some values
are given in Table~X. This values are computed using a calibration
procedure described by Nguyen and Ladd \cite{nguyen-ladd2002}.

@@

\subsubsection{Colloid I/O}

Colloid I/O is controlled via a series of key value pairs. These are:

\inputkey{colloid\_io\_freq}

gives output at this frequency in time steps. Output is also generated
at configuration steps (and measurement steps at the moment).

\inputkey{colloid\_io\_grid} integer vector

sets the colloid I/O grid. It must be accomodated by the current
choice of \texttt{grid}.

\inputkey{colloid\_io\_format\_input}

One of \texttt{ASCII}, \texttt{ASCII\_SERIAL}, \texttt{BINARY}, or
\texttt{BINARY\_SERIAL}

\inputkey{colloid\_io\_format\_output}

For output, only \texttt{ASCII} or \texttt{BINARY} are available.

@@

May be either \texttt{ASCII} or \texttt{BINARY}. Default is \texttt{BINARY}.
Configuration I/O is always in binary. It is recommended to use
binary output for speed and disk space considerations. For a local
sub-domain, the order of the output in the file follows the standard
loop order in the code, i.e., with the $z$-direction running fastest,
then $y$, and with $x$ running slowest. For vector quantities
(anything with more components than a scalar), the vector at each
lattice site appears contiguously.

@@

Individual lattice quantities can use separate I/O grids. This
may be particularly useful for the distribution data, which are
much larger than the other lattice quantities. This means a
finer decomposition may be appropriate. Clearly, the size of the
I/O decomposition cannot exceed the processor decomposition.
See the advice in section \ref{section-advice} for guidelines.

@@

\subsection{Things to Remember and Things to Avoid}
\label{section-advice}

There are a number of simple issues to get right:
\begin{itemize}
\item
Make sure the assertions are switched off (via preprocessor option
\texttt{-DNDEBUG}) and appropriate optimisations
for the current compiler are switched on for production runs (see Makefile).
\item
Do not use too small a local domain size. Typically, 16$^3$ to 32$^3$
cubic local domains will give resaonble scaling. Anything smaller may
be inefficient. The best size may depend on the exact nature of the
calculation and the hardware. Note that this limitation will always
ultimately limit strong scaling.
\item
Don't have diagnsotic output (freq\_statistics) more often than is
necessary. This output requires a global communication which is slow.
\item
Use BINARY output options; ASCII output is slower, and produces larger
files.
\item
Do not use serial I/O. This will have an increasingly high overhead as
the number of processors is increases (it's probably reasonable to
128--256 MPI tasks). This is because each MPI tasks writes its data,
in turn, to a single file; this in inherently serial. To avoid this,
you need to specify the I/O decomposition for the various output
quantities of interest. This splits the system into regular blocks
(in a similar approach to the MPI decomposition), the data for which
are written to separate files in parallel. For example, if the MPI
decomposition is 8\_8\_8 (512 MPI tasks), then a reasonable I/O
decomposition may be 2\_2\_2 or 4\_4\_4. The only way to find out
what works best is to try some tests for the problem size you want to
address.
\item
Don't use the Ewald sum! This simply does not scale.
An alternative algorithm is required.
\end{itemize}
